---
title: "R Notebook"
output: html_notebook
---

# National Household Survey

## Shelter cost


see [2011 National Household Survey: Data tables](http://www12.statcan.gc.ca/nhs-enm/2011/dp-pd/dt-td/Lp-eng.cfm?LANG=E&APATH=3&DETAIL=0&DIM=0&FL=A&FREE=0&GC=0&GID=0&GK=0&GRP=0&PID=0&PRID=0&PTYPE=105277&S=0&SHOWALL=Yes&StartRow=1&SUB=0&Temporal=2013&THEME=98&VID=0&VNAMEE=&VNAMEF=)


```{r}
library(tidyverse)
#library(dplyr)
#library(readr)
#library(tidyr)
#library(ggplot2)

#library(stringr)

#library(xml2)



```


### The Data

> Shelter Cost Groups (11), Household Total Income Groups in 2010 (11), Shelter-cost-to-income Ratio (9), Housing Tenure Including Presence of Mortgage and Subsidized Housing (7) and Household Type (17) for Owner and Tenant Households in Non-farm, Non-reserve Private Dwellings of Canada, Provinces, Territories, Census Metropolitan Areas and Census Agglomerations, 2011 National Household Survey

[99-014-X2011031](http://www12.statcan.gc.ca/nhs-enm/2011/dp-pd/dt-td/Rp-eng.cfm?TABID=4&LANG=E&A=R&APATH=3&DETAIL=0&DIM=0&FL=A&FREE=0&GC=01&GL=-1&GID=1118296&GK=1&GRP=0&O=D&PID=106661&PRID=0&PTYPE=105277&S=0&SHOWALL=0&SUB=0&Temporal=2013&THEME=98&VID=0&VNAMEE=&VNAMEF=&D1=0&D2=0&D3=0&D4=0&D5=0&D6=0)



#### read source file 

A project for another day

** `rsdmx` package **

[rsdmx - Tools for reading SDMX data and metadata in R](https://github.com/opensdmx/rsdmx/wiki#mailinglist_user)


`xml2` package (wormhole back to the tidyverse)
[RStudio Blog: Parse and process XML (and HTML) with xml2](https://blog.rstudio.org/2015/04/21/xml2/)

* [xml2 1.0.0](https://blog.rstudio.org/2016/07/05/xml2-1-0-0/)

* [xml2 v1.1.1](https://www.rdocumentation.org/packages/xml2/versions/1.1.1)

`XML` package
[Read XML data into R](https://www.datacamp.com/community/tutorials/r-data-import-tutorial#gs.RpBzEYA)



---

Use an extract from the IVT file, via Beyond 20/20



```{r  message = FALSE}

# NHS2011data <- read_csv("E:/github/housing-data-visualization-project/data/R3_SCP_Can_PR_CD_CSD_DA_AllGNRs_withSUPP_BC.csv")


#NHS2011data <- read_csv("E:/github/housing-data-visualization-project/data/R3_SCP_Can_PR_CD_CSD_DA_AllGNRs_withSUPP_BC.csv",
#                        col_types = cols(.default = "d",
#                                         Geography = "c"))

NHS2011data <- read_csv("data/NHS_99-014-X2011031_sheltercostratio.csv")

# select only the total columns

NHS2011data_shelter <- NHS2011data %>%
  gather(NHSvar, household.count, -Geography)



```

calculate percentages (x as percentage of region)

```{r}

NHS2011data_shelter %>%
  group_by(Geography) %>%
  mutate(freq = household.count / sum(household.count))

```


