---
  title: "Victoria Population Pyramid"
output:
  html_document: default
html_notebook: default
editor_options:
  chunk_output_type: inline
---
  
  Packages

```{r}

library(tidyverse)
library(feather)
library(plotly)

```


Data

### Read Census data tables from CSV files

```{r}
unzip("98-400-X2016005_English_CSV_data.csv", zipfile = "98-400-X2016005_ENG_CSV.ZIP")
CT_data_Canada <- read_csv("./98-400-X2016005_English_CSV_data.csv")
```


#### decompose geocode variable

> CMACDxx.xx

where

*  CMA  = 3-digit census metropolitan area code
*  CD = 2-digit census division (e.g. municipality) code
*  xx.xx = 2.2-digit dissemination block code 

> If a census tract is split into two or more parts due to a population increase, the number after the decimal point identifies the splits. For example, CT 0042.00 becomes CT 0042.01 and CT 0042.02. If CT 0042.01 is subsequently split, it becomes CT 0042.03 and CT 0042.04. Similarly, if CT 0042.02 is split after CT 0042.01, it becomes CT 0042.05 and CT 0042.06. Any splits occurring after this would be numbered in a similar way, with the next sequential number. This allows users to re-aggregate the splits to the original census tract.

(from ["Naming convention for census tracts"](http://www12.statcan.gc.ca/census-recensement/2016/ref/dict/geo013-eng.cfm))


```{r}

CT_data_Canada <- CT_data_Canada %>%
  mutate(CMA_code = as.integer(substr(`GEO_CODE (POR)`, 1, 3)),
         MUN_code = as.integer(substr(`GEO_CODE (POR)`, 4, 5))
  )


```


rename various variables

```{r}


CT_data_Canada_age <- CT_data_Canada %>%
  rename(age = `DIM: Age (in single years) and average age (127)`,
         pop_total = `Dim: Sex (3): Member ID: [1]: Total - Sex`,
         pop_male = `Dim: Sex (3): Member ID: [2]: Male`,
         pop_female = `Dim: Sex (3): Member ID: [3]: Female`)


```



remove superfluous age categories

```{r}

CT_data_Canada_age_total <- CT_data_Canada_age %>%
  filter(age == "Total - Age")

binned.rows_all <- c("Total - Age", "0 to 14 years", 
                 "0 to 4 years", "5 to 9 years", "10 to 14 years", 
                 "15 to 64 years", "15 to 19 years", 
                 "20 to 24 years", "25 to 29 years", "30 to 34 years", "35 to 39 years",
                 "40 to 44 years", "45 to 49 years", "50 to 54 years", "55 to 59 years",
                 "60 to 64 years", "65 years and over", "65 to 69 years", "70 to 74 years", "75 to 79 years",
                 "80 to 84 years",  "85 years and over", "85 to 89 years",  "90 to 94 years", "95 to 99 years",
                 "Average age")

binned.rows_5yr <- c("0 to 4 years", "5 to 9 years", "10 to 14 years", "15 to 19 years", 
               "20 to 24 years", "25 to 29 years", "30 to 34 years", "35 to 39 years",
               "40 to 44 years", "45 to 49 years", "50 to 54 years", "55 to 59 years",
               "60 to 64 years", "65 to 69 years", "70 to 74 years", "75 to 79 years",
               "80 to 84 years", "85 to 89 years", "90 to 94 years", "95 to 99 years",
               "100 years and over")


CT_data_Canada_age_5yr <- CT_data_Canada_age %>%
  filter(age %in% binned.rows_5yr) 

CT_data_Canada_age_5yr$age <- factor(CT_data_Canada_age_5yr$age, levels = binned.rows_5yr)

levels(CT_data_Canada_age_5yr$age)


CT_data_Canada_age_1yr <- CT_data_Canada_age %>%
  filter(age %in% binned.rows_all)
  
```

---

### Victoria (GEO_CODE = 935XXXX.XX)

Filter CT data


```{r}

CT_data_BC_CMA <- dplyr::inner_join(CT_data_Canada_age_5yr, geoCMA, by = c("CMA_code" = "CMAuid"))


CT_data_age_5yr_Victoria_CMA <- CT_data_Canada_age_5yr %>% 
  filter(CMA_code == "935")

CT_data__age_5yr_Victoria_mun <- CT_data_age_5yr_Victoria_CMA %>% 
  filter(MUN_code == "00")

```


write files

```{r}

# csv
write_csv(CT_data__age_5yr_Victoria_mun, 
          "./CT_data__age_5yr_Victoria_mun.csv")


# feather
write_feather(CT_data__age_5yr_Victoria_mun, 
          "./CT_data__age_5yr_Victoria_mun.feather")



```
















```{r}
PopData_source <- read_feather("./CT_data__age_5yr_Victoria_mun.feather")
```



Summarize for Victoria City

```{r}

final.pop.raw <- PopData_source %>%
  group_by(age) %>%
  summarise(male = sum(pop_male),
            female = sum(pop_female)) %>%
  gather("male", "female", key = "sex", value = "pop")


final.pop.raw


# Normalize to age bins as a share of the total population in that year.

final.pop <- final.pop.raw %>%
  #  group_by(sex) %>%
  mutate(poppct = pop / sum(pop)) %>%
  ungroup() %>%
  mutate("neighbourhood" = "Victoria (City)")


final.pop

```


```{r}

ggplot(final.pop, aes(x = age, fill = sex)) +
  geom_bar(aes(y = poppct), stat = "identity", position = "identity",
           color = "gray10", size = 0.1,
           data = filter(final.pop, sex == "male")) +
  geom_bar(aes(y = -1 * poppct), stat = "identity", position = "identity",
           color = "gray10", size = 0.1,
           data = filter(final.pop, sex == "female")) +
  #  geom_text(aes(label = year), y = 0.06, x = 19.5, family = "minion",
  #            size = 3) +
  #  facet_grid(geo ~ year, scales = "fixed") +
  coord_flip() +
  scale_fill_brewer(palette = "Pastel1") +
  labs(x = NULL, y = NULL, title = NULL, fill = NULL) +
  theme_bw()

```

Make charts using plotly

```{r}
final.pop <- final.pop %>%
  mutate(abspop = pop) %>%
  mutate(pop = ifelse(sex == "male", pop * -1, pop))

plot_ly(final.pop, x = ~pop, y = ~age, color = ~sex, type = 'bar', orientation = 'h',
        hoverinfo = 'y+text+name', text = ~pop, colors = c('salmon', 'lightblue')) %>%
  layout(bargap = 0.1, barmode = 'overlay',
         xaxis = list(tickmode = 'array'#,
                      # tickvals = c(-10000, -5000, 0, 5000, 10000),
                      # ticktext = c('10k', '5k', '0', '5k', '10k')
         ),
         title = 'Projected population structure of China, 2050')
```



```{r}
library(grid)
library(gridExtra)
library(cancensus)

# regions <- list_census_regions('CA16') %>% filter(CMA_UID=="59933",level=="CSD") %>% top_n(6,pop)

# plots <- lapply(seq(1,nrow(regions)),function(i,rs){
#   r=rs[i,]
#   plot_data <- get_age_data('CA16',as_census_region_list(r))
# 
#  p <- ggplot(plot_data, aes(x = Age, y = Population, fill = Gender)) + 
#   geom_bar(stat = "identity") +
#   age_pyramid_styling +
#   #scale_fill_discrete(guides=FALSE) +
#   labs(title=paste0(r$name," (",r$municipal_status,")"),x="",y="",caption="") +
#   theme(legend.position = "none")
# },rs=regions)

# grid.arrange(
#   grobs=plots,
#   ncol = 2,
#   top = 'Population by Age for Metro Vancouver Municipalities',
#   bottom="Canada Census 2016 via cancensus & CensusMapper.ca"
#   )
```
